Wall-to-wall mapping of peat depth from Lidar terrain and airborne radiometrics in Norwegian landscapes

### 2024-10-30 ms/ms.Rmd
ms: minor edits - model performance

  1. only DMK peat depth class (2 binary indicator variables from 3 levels)
  2. only terrain (21 continuous variables)
  3. terrain and DMK peat depth class (23 variables)
  4. radiometric and terrain predictors (25 continuous variables)
  5. all predictors (26 variables)

@@

While the primary aim of the DSM was to predict peat depth within peatlands (where peat > 30 cm), we also performed a minimal evaluation of our ability to estimate peat depth outside of mapped peatland.
In other words, we tested if predictions of deep peat indicated unmapped peatland occurrence in the wider landscape.
It is practically difficult to obtain suitable data for this problem, because locations with non-zero peat depth outside of mapped peatland are unknown and relatively rare.
Therefore, we made do with the same data that we used for the peat depth predictions, which included a fraction of locations not mapped as peatland.
We used the same spatial cross-validation as for the peat depth predictions, but this time we trained the models only on locations mapped as peatland, and evaluated them only on locations not mapped as peatland.
In this way, we simulated the situation where a model trained on known peatland is used to predict peat depth outside the known peatland.
We evaluated models with radiometric and terrain predictors, or terrain predictors alone (since DMK peat depth class is always undefined outside known peatlands).
From the cross-validation we quantified *Root Mean Squared Error* and *Mean Absolute Error*, and we tested the strength of the correlation between predicted and observed peat depth.

>>>>

  1. only DMK peat depth class (2 binary indicator variables from 3 levels)
  2. only terrain (21 continuous variables)
  3. terrain and DMK peat depth class (23 variables)
  4. terrain and radiometric predictors (25 continuous variables)
  5. all predictors (26 variables)

@@

While the primary aim of the DSM was to predict peat depth within peatlands (where peat > 30 cm), we also performed a minimal evaluation of our ability to estimate peat depth outside of mapped peatland.
In other words, we tested if predictions of deep peat indicated unmapped peatland occurrence in the wider landscape.
It is practically difficult to obtain suitable data for this problem, because locations with non-zero peat depth outside of mapped peatland are unknown and relatively rare.
Therefore, we made do with the same data that we used for the peat depth predictions, which included a fraction of locations not mapped as peatland.
We used the same spatial cross-validation folds, but this time we trained the models only on locations mapped as peatland (in the training folds), and evaluated them only on locations not mapped as peatland (in the test fold).
In this way, we simulated the situation where a model trained on known peatland is used to predict peat depth outside the known peatland.
Since DMK peat depth class is always undefined outside known peatlands, we evaluated two predictor configurations: terrain predictors alone or terrain and radiometric predictors.
From the cross-validation we quantified *Root Mean Squared Error* and *Mean Absolute Error*, and we tested the strength of the correlation between predicted and observed peat depth.

### 2024-10-30 ms/ms.tex
ms: minor edits - model performance

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  only DMK peat depth class (2 binary indicator variables from 3 levels)
\item
  only terrain (21 continuous variables)
\item
  terrain and DMK peat depth class (23 variables)
\item
  radiometric and terrain predictors (25 continuous variables)
\item
  all predictors (26 variables)
\end{enumerate}

@@

While the primary aim of the DSM was to predict peat depth within peatlands (where peat \textgreater{} 30 cm), we also performed a minimal evaluation of our ability to estimate peat depth outside of mapped peatland.
In other words, we tested if predictions of deep peat indicated unmapped peatland occurrence in the wider landscape.
It is practically difficult to obtain suitable data for this problem, because locations with non-zero peat depth outside of mapped peatland are unknown and relatively rare.
Therefore, we made do with the same data that we used for the peat depth predictions, which included a fraction of locations not mapped as peatland.
We used the same spatial cross-validation as for the peat depth predictions, but this time we trained the models only on locations mapped as peatland, and evaluated them only on locations not mapped as peatland.
In this way, we simulated the situation where a model trained on known peatland is used to predict peat depth outside the known peatland.
We evaluated models with radiometric and terrain predictors, or terrain predictors alone (since DMK peat depth class is always undefined outside known peatlands).
From the cross-validation we quantified \emph{Root Mean Squared Error} and \emph{Mean Absolute Error}, and we tested the strength of the correlation between predicted and observed peat depth.

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  only DMK peat depth class (2 binary indicator variables from 3 levels)
\item
  only terrain (21 continuous variables)
\item
  terrain and DMK peat depth class (23 variables)
\item
  terrain and radiometric predictors (25 continuous variables)
\item
  all predictors (26 variables)
\end{enumerate}

@@

While the primary aim of the DSM was to predict peat depth within peatlands (where peat \textgreater{} 30 cm), we also performed a minimal evaluation of our ability to estimate peat depth outside of mapped peatland.
In other words, we tested if predictions of deep peat indicated unmapped peatland occurrence in the wider landscape.
It is practically difficult to obtain suitable data for this problem, because locations with non-zero peat depth outside of mapped peatland are unknown and relatively rare.
Therefore, we made do with the same data that we used for the peat depth predictions, which included a fraction of locations not mapped as peatland.
We used the same spatial cross-validation folds, but this time we trained the models only on locations mapped as peatland (in the training folds), and evaluated them only on locations not mapped as peatland (in the test fold).
In this way, we simulated the situation where a model trained on known peatland is used to predict peat depth outside the known peatland.
Since DMK peat depth class is always undefined outside known peatlands, we evaluated two predictor configurations: terrain predictors alone or terrain and radiometric predictors.
From the cross-validation we quantified \emph{Root Mean Squared Error} and \emph{Mean Absolute Error}, and we tested the strength of the correlation between predicted and observed peat depth.
